The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-12-11 manuscript.Rmd
[fix #88] add mention of low level of digitization on West coast
* for CAS website reports collection has 2.5 million but < 165k on iDigBio
* 0 records for LACM
[fix #83] finish part about BLM/MMS contributions

Another indication of the inadequacy of the sampling is the relationship between the number of collected specimens and the number of species recorded through time (Fig. \@ref(fig:idigbio-through-time)). Explorations in the late nineteenth century led to an exponential growth in the knowledge of the fauna, but collecting then tapered off, until the 1970s when intensive collecting efforts led to an increase in the recording of new taxa (Fig. \@ref(fig:idigbio-through-time)). Since then, the slowed pace of collection activitity (and therefore a reduction in the number of specimens added) has been accompanied by a slow-down in the discovery of new records: `r n_spp_comparison$n_spp_2017_2000` species have been documented for the first time since 2000 (`r format_output(n_spp_comparison$n_spp_2017_2000/18)` species.year^-1^), while `r n_spp_comparison$n_spp_1970_1960` were documented between 1960 and 1970 (`r format_output(n_spp_comparison$n_spp_1970_1960/11)` species.year^-1^). The strong correlations between (1) the number of samples collected and the species richness (Fig. \@ref(fig:correlation-sampling)), and (2) the number of specimens collected per year and the number of species recorded for the first time (Fig. \@ref(fig:idigbio-through-time)C), suggest that any new sampling effort will lead to new observations and the discovery of new species.

@@

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio, we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for those that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, is needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems.

>>>>

Another indication of the inadequacy of the sampling is the relationship between the number of collected specimens and the number of species recorded through time (Fig. \@ref(fig:idigbio-through-time)). Explorations in the late nineteenth century led to an exponential growth in the knowledge of the fauna, but collecting then tapered off, until the 1970s when intensive collecting efforts led to an increase in the recording of new taxa (Fig. \@ref(fig:idigbio-through-time)). Since then, the slowed pace of collection activitity (and therefore a reduction in the number of specimens added) has been accompanied by a slow-down in the discovery of new records: `r n_spp_comparison$n_spp_2017_2000` species have been documented for the first time since 2000 (`r format_output(n_spp_comparison$n_spp_2017_2000/18)` species.year^-1^), while `r n_spp_comparison$n_spp_1970_1960` were documented between 1960 and 1970 (`r format_output(n_spp_comparison$n_spp_1970_1960/11)` species.year^-1^).

```{r idigbio-mms}
prop_mms_75_85 <- (idigbio_mms_summary_by_year %>%
                   dplyr::filter(year >= 1975 & year <= 1985) %>%
                   dplyr::pull(n) %>%
                   sum()) / dplyr::pull(idigbio_mms_summary_by_year, n)
```

Biodiversity surveys conducted under the auspices of the Bureau of Land Management (BLM) and the Minerals Management Service (MMS) have contributed significantly to our knowledge of the geographic distribution of marine invertebrates in US waters. In our dataset, we identified at least `r format(sum(idigbio_mms_summary$n), big.mark=",")` records (or `r sum(idigbio_mms_summary_by_year$n)/nrow(idigbio_us_inverts_records_all)*100`% of all US marine invertebrates in iDigBio) from these surveys, `r format_output(prop_mms_75_85 * 100)`% of which were collected between 1975 and 1985. Of the `r idigbio_mms_records %>% filter(rank %in% c("Species", "Subspecies")) %>% summarize(n = n_distinct(worms_valid_name))` species collected as part of these surveys, `r nrow(idigbio_mms_species_only)` have not been collected outside of these surveys.

The strong correlations between (1) the number of samples collected and the species richness (Fig. \@ref(fig:correlation-sampling)), and (2) the number of specimens collected per year and the number of species recorded for the first time (Fig. \@ref(fig:idigbio-through-time)C), suggest that any new sampling effort will lead to new observations and the discovery of new species.

@@

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio, we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for those that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, is needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems. Similarly, only a small fraction of the marine invertebrates records from instiutions on the Pacific have been digitized.
